The Spatial and Temporal Domains of Modern Ecology

### 2017-12-19 paper/main.tex
Edited main text through difference magnitudes. Made more edits to figure S3 code, fixing too wide boxes in paleo duration. Added return to tops in a couple other vignettes

Contrasting the two spatial dimensions against one another (for all observations) shows a primary concentration of observations with 10 cm$^2$ to 100 m$^2$ resolution with extents ranging from just over 1,000 to nearly 1,000,000 ha (Fig. 2C). The second-most prominent concentration consists of higher resolution (1 cm$^2$-1 m$^2$), smaller extent (10-1,000 ha) observations, beneath which lies a third and fainter concentration of 1-1,000 cm$^2$ resolution, 1000 m$^2$ to $<$10 ha. These three concentrations suggest a tendency for observational extent to increase with resolution, which is further evident in the median domain values (and kernel densities; Fig. S2 in SI) of automated (0.01 m$^$ resolution, 100 ha extent), field (0.1-1 m$^2$ resolution, 1,000-10,000 ha extent), and remote (1,000 m$^2$ resolution, 1-10 million ha extent) observations.  Paleo-reconstructions were an outlier from this relationship, having very fine median resolution (0.01 m$^2$) but large extent (1 million ha), a result that likely reflects the very low sample size for this observational type.

@@

\noindent \textbf{\emph{Differences between actual and ostensible scales}}\\
For our final analysis of observational domains, we evaluated the differences between the scales represented by the extent and duration dimensions and the scales that ecological observations actually cover. To make this assessment, we log$_{10}$ transformed and then subtracted the values of i) actual extent from extent and ii) actual duration from duration, yielding the magnitude of difference between each pair of dimensions for each observation. We plotted these values (summarized in box plots) against their corresponding extent/duration values to evaluate whether these differences vary with scale (Fig. 3). This analysis showed that most (81\%) of the assessed observations had actual extents of $\leq$1 ha, which on average were 4 to nearly 8 orders of magnitude smaller than the primary measure of extent (Fig 3A). These differences declined linearily with actual extent, becoming approximately equal for the 5\% of observations having $\geq$100,000 ha actual extent. The actual duration of 64\% of observations was $<$1 day, which on average was 3 to nearly 9 orders of magnitude shorter than the time span covered by the conventional metric of duration (Fig. 3B). As with extent, the average magnitude of difference declined steadily with actual duration, becoming effectively equal for the 16\% of observations with $\geq$1 month of actual duration.

%\noindent \textbf{\emph{Observational methods}}\\
%We classified the method used to collect each observation into several broad categories, which were field methods (manual \emph{in situ} data collection), automated (\emph{in situ}) sensing, remote sensing, other geographic data, and paleo-reconstruction approaches. Field methods were used for 80\% of observations, automated sensing for 12.4\%, remote sensing for 6.4\%, and paleo-reconstruction and other geographic data each for less than 1\% each.

>>>>

Contrasting the two spatial dimensions against one another (for all observations) shows a primary concentration of observations with 10 cm$^2$ to 100 m$^2$ resolution with extents ranging from just over 1,000 to nearly 1,000,000 ha (Fig. 2C). The second-most prominent concentration consists of higher resolution (1 cm$^2$-1 m$^2$), smaller extent (10-1,000 ha) observations, beneath which lies a third and fainter concentration of 1-1,000 cm$^2$ resolution, 1000 m$^2$ to $<$10 ha. These three concentrations suggest a tendency for observational extent to increase with resolution, which is further evident in the median domain values (and kernel densities; Fig. S2 in SI) of automated (0.01 m$^2$ resolution, 100 ha extent), field (0.1-1 m$^2$ resolution, 1,000-10,000 ha extent), and remote (1,000 m$^2$ resolution, 1-10 million ha extent) observations.  Paleo-reconstructions were an outlier from this relationship, having very fine median resolution (0.01 m$^2$) but large extent (1 million ha), a result that likely reflects the very low sample size for this observational type.

@@

\noindent \textbf{\emph{Differences between actual and ostensible scales}}\\
For our final analysis of observational domains, we evaluated the differences between the scales represented by the extent and duration dimensions and the scales that ecological observations actually cover. To make this assessment, we log$_{10}$ transformed and then subtracted the values of i) actual extent from extent and ii) actual duration from duration, yielding the magnitude of difference between each pair of dimensions for each observation. We plotted these values (summarized in box plots) against their corresponding extent/duration values to evaluate whether these differences varied with scale (Fig. 3). Extent was on average 5.6 orders of magnitude larger than actual extent, and this difference increased with extent, reaching a maximum of 8 orders of magnitude between 100 million and 1 billion ha of extent (Fig. 3A). The difference fell to 3 orders of magnitude at 10 billion ha, a domain that was covered by $<$2\% of observations that were primarily collected with remote sensing. Remote observations had the smallest mean difference magnitude (1.9), whereas differences for the other three methods averaged at least 5.7 orders of magnitude larger (Fig. S3 in SI).

The magnitude of difference between the duration and actual duration of observations was somewhat smaller, averaging 3.4 across all observations, ranging from just under 2 for the shortest durations (hour-day) to over 4 for observations lasting 1 decade to 1 century (Fig. 3B). As with extent, the difference fell substantially for the longest durations (century to 10,000 years), as these domains were covered by paleo-reconstructions (Fig. S3 in SI), which have effectively no difference between duration and actual duration because the coring techniques they are typically based on capture continuous temporal records. The mean difference magnitudes for the other three observing methods ranged from just over 3 for field and automated sensing methods to nearly 6 for remote observations (Fig. S3 in SI).
